---
title: "Final Project"
author: "Kelly Cook"
output: html_document
---

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

<style>
div.gray { background-color:#9CD3A9; border-radius: 5px; padding: 20px;}
</style>
<div class = "gray">

## First Responders Mental Health Awareness Survey

<style>
div.w { background-color:#ffffff; border-radius: 5px; padding: 20px;}
</style>
<div class = "w">

```{r setup chunk, setup, warning=FALSE, message=FALSE, fig.align='center', include=FALSE}
require("knitr")
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.align='center', echo = TRUE)
library(pacman)
pacman::p_load("corrplot", "tidyverse", "modelr", "dplyr", "GGally", "patchwork", "jtools", "skimr", "descr", "gmodels")
library(corrplot)
```

### Survey Data Analysis 

#### Background

This survey is considered a "First Responders Mental Health Awareness Survey" or a Wellness and Support Questionnaire. 
It's designed to give the administration a picture of how their department is doing based on their members' self-reporting. 
The survey is broken up into sections measuring different areas of life first responders are typically most affected in. 

#### Cleaning the data

Cleaning the raw data was an extencive work. This particular data sets department membership has 545 sworn officers, 119 civilian members (non-officers). 324 responded (but not to every question)-approximately 48.80% response rate.There is an .Rproj in this folder that contains the primary cleaning.
The data will be divided in five color coded sections for convinience. 

### Data overview

```{r message=FALSE, warning=FALSE, results='hide'}
df <- read.csv("./Police_clean.cvs")
skim(df)
glimpse(df)
df <- na.omit(df)
df$Position <- factor(df$Position) 
levels(df$Position) <- c("Sworn", "Disp")
df$how.many.years <- factor(df$how.many.years)
levels(df$how.many.years) <- c("1-6", "7-13", "14-20", "20+")
```

The survey was administered to Sworn Officer and Law Enforcement Professional (Dispatch or other Civilian Department Member). Which if found in the data as Position =  "Sworn" and "Disp"

### Distribution

```{r message=FALSE, warning=FALSE}
prop.table(table(df$Position))
```

  - Meaning 79.2% Sworn Officer and 20.8% Law Enforcement Professional (Dispatch or other Civilian Department Member)

#### Distribution according to time of service 

```{r}
CrossTable(df$how.many.years,df$Position,prop.r=FALSE,prop.t=FALSE,prop.chisq=FALSE,format="SPSS")
```

Another way to visualice this information is with a graph. 
```{r Map fig1, results="hide", message=FALSE, warning=FALSE,fig.align = 'center', fig.cap="**Figure 1a.** *Distribution of the subject participating in the survey according to their position and time of service.*"}
ggplot(df, aes(x= how.many.years, fill= factor(how.many.years))) +
  labs(x= "Years of Service", fill= "Years of Service") +
  geom_bar() +
  facet_grid(~Position)
```


#### Note:

  - 2. Due to the steps are same for each secction. The Yelow section will serve as refference for the steps used to analysed the data.
  
  - 1. Each section contains multiple stadistical models. To reduced the size of this document not all models will be graphed.

<br>
</div>
___

<style>
div.yellow { background-color:#ffff7a; border-radius: 5px; padding: 20px;}
</style>
<div class = "yellow">
### Yellow

#### Programs Training or service already in place or potentially being considered

```{r message=FALSE, warning=FALSE}
dat <- read.csv("./Programs_Yellow1.cvs")
dat <- na.omit(dat)
dat$Years <- factor(dat$Years)
levels(dat$Years) <- c("1-6 years", "7-13 years", "14-20 years", "20+ years")
```

  - This section is diveded in EPA questions, Peer support, In job Trainnings, Personal Workshops. 

  - The procces of the analysis will be divided in 5 teps for your convinience. 
  
  - The steps will be well explained and shown in the first two sections. (You can request changes, or request to show only your preffered steps)
  
  _ The steps are reproduceble for each section of the data. For convinience I created separete .cvs folders to access each section indivualy if needed.  

### EPA

#### Step1
#### There are a couple of ways to visualize the survey data.

  - Simple graphs 
```{r Map fig2, results="hide", message=FALSE, warning=FALSE,fig.align = 'center', fig.cap="**Figure 2a.**Do you know what EAP services are available to you? Answers **"}
ggplot(dat, aes(x= whatE, fill= factor(whatE))) +
  labs(x= "answer", title = "Do you know what EAP services are available to you?") +
  scale_fill_discrete(name = "Questions",
labels = c("not at all", "a little bit", "moderately", "quite a bit", "extremely")) +
  geom_bar() +
  theme_bw() +
  facet_grid(~Years) +
  theme(strip.text.x = element_text(face= "bold"))
```

  - Tables

```{r message=FALSE, warning=FALSE}
#How Likely are you to use EAP Services
CrossTable(dat$Years,dat$likeE,prop.r=FALSE,prop.t=FALSE,prop.chisq=FALSE,format="SPSS")
```
             
  - The table allows to have a numeric value for the responce "not at all" 11%, "a little bit" 37.5%, "moderately" 37.8% , "quite a bit" 11.6%, "extremely" 1.7%

  - Both table and graph can be done for each individual question. (I am not sure what would you prefferer?)

#### Step2
#### Are this questions related?

```{r Map fig3, results="hide", message=FALSE, warning=FALSE,fig.align = 'center', fig.cap="**Figure 2b.** *Correlation grph. EPA Qiestions*"}
dat1 <- select(dat, whatE, howE, likeE)
dat1c <- cor(dat1)
corrplot.mixed(dat1c)
```
We can observe a positive and negative correlation. Models would depend in desired outcome of the data

#### Step4
#### Finding the best model

```{r message=FALSE, warning=FALSE}
mod1 <- lm(formula = likeE ~ whatE, data = dat)
mod2 <- lm(formula = likeE ~ howE, data = dat)
summary(mod2)#whatE is a better predictor to likeE
mod1mse <- mean(residuals(mod1)^2)
mod2mse <- mean(residuals(mod2)^2)
su2<- summary((mod2))$r.squared
su1<- summary((mod1))$r.squared
mod1mse ; mod2mse ; su1 ; su2
```
whatE seems to be a beeter predictore for likeE. However, it explains only 10% of the varience.  

#### Step5
#### Graphing models + predicctions 
```{r Map fig4, results="hide", message=FALSE, warning=FALSE,fig.align = 'center', fig.cap="**Figure 2c.**Likely to usu EPA with respect What is EPA**"}
effect_plot(mod1, pred = whatE, plot.points = TRUE, jitter = 0)
```

### Peer support

```{r Map fig5, results="hide", message=FALSE, warning=FALSE,fig.align = 'center', fig.cap="**Figure 3a.** *Correlation grph. Peer Support Qiestions*"}
dat3 <- select(dat, whatP, whoP, howP, likeP, comfP)
dat3c <- cor(dat3)
corrplot.mixed(dat3c)
```

Best model previusly determined
```{r message=FALSE, warning=FALSE}
mod4 <- lm(data=dat3, formula = whoP ~ howP)
su4<- summary((mod4))$r.squared
mod4mse <- mean(residuals(mod4)^2)
mod4mse ; su4 
```

```{r Map fig6, results="hide", message=FALSE, warning=FALSE,fig.align = 'center', fig.cap="**Figure 3b.** *Correlation grph. Who is you peer support team? with respect How to access Peer support?*"}
effect_plot(mod4, pred = howP, plot.points = TRUE, jitter = 0)
```

### In job Trainnings

```{r Map fig7, results="hide", message=FALSE, warning=FALSE,fig.align = 'center', fig.cap="**Figure 4a.** *Correlation grph. In job Trainnings*"}
dat4 <- select(dat, skillsT, partT, shift, wellT)
dat4c <- cor(dat4)
corrplot.mixed(dat4c)
```

```{r message=FALSE, warning=FALSE}
mod5 <- lm(data=dat4, formula = skillsT ~ partT)
su5<- summary((mod5))$r.squared
mod5mse <- mean(residuals(mod5)^2)
mod5mse ; su5
```

### Personal Workshops

```{r Map fig8, results="hide", message=FALSE, warning=FALSE,fig.align = 'center', fig.cap="**Figure 5a.** *Correlation grph. Personal Workshops *"}
dat5 <- select(dat, diet, exerc, fam, suppG, marry, finance, mind, resi, reimb)
dat5c <- cor(dat5)
corrplot.mixed(dat5c)
```
           
  - Correlations seems to be stronger in this section
  
```{r Map fig9, results="hide", message=FALSE, warning=FALSE,fig.align = 'center', fig.cap="**Figure 5b.** *Correlation grph. Personal Workshops *"}
ggpairs(dat5)
```

  - Another way to see correlations with grphs instead of color coded. 
  
```{r message=FALSE, warning=FALSE}
mod6 <- lm(data=dat5, formula = diet ~ exerc)
su6<- summary((mod6))$r.squared
mod6mse <- mean(residuals(mod6)^2)
mod6mse ; su6
```

```{r Map fig9.1, results="hide", message=FALSE, warning=FALSE,fig.align = 'center', fig.cap="**Figure 5c.** *Correlation grph. Predicting Diet workshop attendace based on Exercise workshop attendace *"}
effect_plot(mod6, pred = exerc, plot.points = TRUE, jitter = 0)
```
<br>
</div>
___

<style>
div.blu { background-color:#0085ba; border-radius: 5px; padding: 20px;}
</style>
<div class = "blu">
#### Blue 
### Spiritual Impact

```{r message=FALSE, warning=FALSE}
df <- read.csv("./Programs_Blue.cvs")
df$how.many.years <- factor(df$how.many.years)
levels(df$how.many.years) <- c("1-6", "7-13", "14-20", "20+")
dfa <- subset(df, select = c(3:8))
```

```{r message=FALSE, warning=FALSE}
CrossTable(df$how.many.years,df$extent.spiritual,prop.r=FALSE,prop.t=FALSE,prop.chisq=FALSE,format="SPSS")
```

```{r Map fig10, results="hide", message=FALSE, warning=FALSE,fig.align = 'center', fig.cap="**Figure 10a.** *Correlation grph. Spiritual Impact Questions *"}
dfac <- cor(dfa)
corrplot.mixed(dfac)
```

```{r message=FALSE, warning=FALSE}
moddf <- lm(data=df, formula = spritual.help ~ spiritual.tolerate)
sudf<- summary((moddf))$r.squared
moddfmse <- mean(residuals(moddf)^2)
moddfmse ; sudf
```
  
  - Explains 85% of the varience
  
```{r Map fig11, results="hide", message=FALSE, warning=FALSE,fig.align = 'center', fig.cap="**Figure 10b.** *Correlation grph. Stress Tolerance based on the Speritual help *"}
effect_plot(moddf, pred = spiritual.tolerate, plot.points = TRUE, jitter = 0)
```
<br>
</div>
___

<style>
div.red { background-color:#ff5d3d; border-radius: 5px; padding: 20px;}
</style>
<div class = "red">

#### Red 
### PTSD, Depression, Anxiety and Suicidal Measures

```{r message=FALSE, warning=FALSE}
dfr <- read.csv("./Programs_Red.cvs")
dfr$how.many.years <- factor(dfr$how.many.years)
levels(dfr$how.many.years) <- c("1-6", "7-13", "14-20", "20+")
```

```{r Map fig12, results="hide", message=FALSE, warning=FALSE,fig.align = 'center', fig.cap="**Figure 12a.** *Feeling down, depressed, or hopeless? Answer rate *"}

ggplot(dfr, aes(x= feeling.depressed , fill= factor(feeling.depressed))) +
  labs(x= "answer", title = "Feeling down, depressed, or hopeless?") +
  scale_fill_discrete(name = "Answers",
labels = c("not at all", "a little bit", "moderately", "quite a bit", "extremely")) +
  geom_bar() +
  theme_bw() +
  facet_grid(~how.many.years) +
  theme(strip.text.x = element_text(face= "bold"))
```
```{r message=FALSE, warning=FALSE}
CrossTable(dfr$how.many.years,dfr$feeling.depressed,prop.r=FALSE,prop.t=FALSE,prop.chisq=FALSE,format="SPSS")
```
<br>
</div>
___

<style>
div.oran { background-color:#ff9b3d; border-radius: 5px; padding: 20px;}
</style>
<div class = "oran">
#### Orange 
### Barriers to getting help

```{r message=FALSE, warning=FALSE}
dfo <- read.csv("./Programs_Orange.cvs")
dfo$how.many.years <- factor(dfo$how.many.years)
levels(dfo$how.many.years) <- c("1-6", "7-13", "14-20", "20+")
```

```{r message=FALSE, warning=FALSE}
CrossTable(dfo$how.many.years,dfo$I.might.need,prop.r=FALSE,prop.t=FALSE,prop.chisq=FALSE,format="SPSS")
```

  - Almost 40% of personal has wonder if they need mental health
  
```{r Map fig13, results="hide", message=FALSE, warning=FALSE,fig.align = 'center', fig.cap="**Figure 13a.** *At times I have wondered if I need mental health treatment? Answer rate *"}
ggplot(dfo, aes(x= I.might.need , fill= factor(I.might.need))) +
  labs(x= "answer", title = "At times I have wondered if I need mental health treatment?") +
  scale_fill_discrete(name = "Answers",
labels = c("not at all", "a little bit", "moderately", "quite a bit", "extremely")) +
  geom_bar() +
  theme_bw() +
  facet_grid(~how.many.years) +
  theme(strip.text.x = element_text(face= "bold"))
```
<br>
</div>
___

<style>
div.pur { background-color:#ad3bff; border-radius: 5px; padding: 20px;}
</style>
<div class = "pur">
#### Purple 
### Coping

```{r message=FALSE, warning=FALSE}
dfp <- read.csv("./Programs_Purple.cvs")
dfp$how.many.years <- factor(dfp$how.many.years)
levels(dfp$how.many.years) <- c("1-6", "7-13", "14-20", "20+")
```

```{r message=FALSE, warning=FALSE}
CrossTable(dfp$how.many.years,dfp$how.often.alcohol,prop.r=FALSE,prop.t=FALSE,prop.chisq=FALSE,format="SPSS")
```

```{r Map fig14, results="hide", message=FALSE, warning=FALSE,fig.align = 'center', fig.cap="**Figure 14a.** *How often do you drink? Answer rate *"}
names(dfp)
ggplot(dfp, aes(x= how.often.alcohol , fill= factor(how.often.alcohol))) +
  labs(x= "answer", title = "How often do you drink?") +
  scale_fill_discrete(name = "Answers",
labels = c("not at all", "a little bit", "moderately", "quite a bit", "extremely")) +
  geom_bar() +
  theme_bw() +
  facet_grid(~how.many.years) +
  theme(strip.text.x = element_text(face= "bold"))
```
<br>
</div>
___




